Title

### 2012-12-05 inc/photocathode/plasmon/plasmon.tex
added diagram describing the rotated geometry
more work on the plasmonic emission discussion

\begin{figure}
  \centering
  \input{power_seq}
  \caption{}
  \label{fig:pape}
\end{figure}

% Content taken from 2010 ANL report
Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
These two properties are known characteristics of plasmon-assisted photoemission.

As seen in \ref{fig:pape}, at low incident laser intensities, below about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
Moreover, both emissions have similar momentum distributions \fixme{(images A and E)}, again indicating a similar emission process.
The only difference is that the emission from the nano-patterned region is 3-4 times more efficient.
This is readily explained by the fact that the periodic surface modulation results in a local enhancement of the $\sim$1MV/m DC gun field at the peaks of the modulation which, through the Schottky effect, then presents a reduced barrier for field emission.

Above $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
The observed distinct changes in the momentum distributions \fixme{(images B and C)} in this laser intensity range support this argument.
In \fixme{image B}, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves). 
This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \fixme{image D} and by the cut-off in the detected electron signal (YAG fluorescence).

@@

>>>>

\begin{figure}
  \centering
  \input{power_seq}
  \caption{
    Photoemission characteristics of laser-driven plasmons on a one-dimensional periodic gold surface.
    Top: Electron emission strength as a function of incident 523nm femtosecond laser pulse intensity for coherently-driven plasmons (top curve)
      % with parabolic fit to initial dependence (dashed curve)
      and comparative emission from a plane gold surface under the same irradiantion conditions.
    Bottom: Two-dimensional electron momentum distributions for plasmon-assisted emission (A-D) and two-photon field emission from gold (E) at the laser intensities indicated on the top graph.
    The circular aperture visible in B-D is the 6mm-diameter bore of the magnetic focusing lens.
    The orientation of the grating lines of the periodic gold surface with respect to the emission in A-D is parallel to the broadening seen in B-D.
  }
  \label{fig:pape}
\end{figure}

% Content taken from 2010 ANL report
Note that for this measurement, relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
The scintillator and CCD camera detection scheme was also shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
As a result, the detection scheme will accurately determine any laser power law dependence to the electron emission from the photocathode.

Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
These two properties are known characteristics of plasmon-assisted photoemission.

As seen in \ref{fig:pape}, at low incident laser intensities, below about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
Moreover, both emissions have similar momentum distributions (\ref{fig:pape} A and E), again indicating a similar emission process.
The only difference is that the emission from the nano-patterned region is 3-4 times more efficient.
This is readily explained by the fact that the periodic surface modulation results in a local enhancement of the $\sim$1MV/m DC gun field at the peaks of the modulation which, through the Schottky effect, then presents a reduced barrier for field emission.

Above $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
The observed distinct changes in the momentum distributions (\ref{fig:pape} B and C) in this laser intensity range support this argument.
In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves). 
This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \ref{fig:pape} D and by the cut-off in the detected electron signal (YAG fluorescence).

@@

In principle this unexpected increase in transverse momentum could be caused by either of two effects.
The first is the divergence of the plasmon's electric field ($\partial E_{plasmon} / \partial x$) parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
The second is the one dimensional local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface (parallel to the grating vector $k_G$); this is commonly known as the ``lightning rod'' effect.

Given the geometry of the previous measurement, notably the colinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.

\begin{figure}
  \centering
  \input{rotated}
  \caption{
    Comparison of the colinear (above) and rotated (below) plasmonic excitation geometries.
    The vector diagrams (left) are related to the corresponding momentum distribution image (right).
    That the distribution is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed) is evidence that the grating is the cause of the increased momentum distribution.
  }
  \label{fig:rotated}
\end{figure}

\ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
For this rotation, the laser angle must be increased from $39^{\circ}$ to $47^{\circ}$.

The results are shown to the right in \ref{fig:rotated}.
As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.

This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
As that geometry does not employ a patterned surface, it would is not subject to such local field enhancements of the accelerating electric field.
